---
title: "psych710_final"
output: html_document
date: "2024-04-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background/Hypothesis
The dataset that we have chosen to use is from Sparkman et al. 2022 article, "Americans experience a false social reality by underestimating popular climate policy support by nearly half." Our research question is How do socioeconomic factors predict green policy ideologies? The data that we have comes from a survey census (Ipsos eNation Omnibus) that utilizes questions for misperception for others' thoughts and behaviors by concerning climate change and support for mitigation policies, comparing with the ground truth of support of climate policies (from Yale's program on Climate Change Communication). The outcome variable, levels of support/more concern for climate policies, were measured twice with different wording. Overall, we predict that individuals' socioeconomic statuses will affect the tendency of changing levels of support/more concern for climate policies. 

## Proposed Analysis
For sure, we will be using a Bayesian statistic to predict climate policy support from socioeconomic factors which takes into account the repeated measures between the two climate change/policy-related questions with different wording that these surveys were sent out. In our analysis, we will create two composite scores of climate policy support and look for which factor(s) would significantly affect their tendency for climate policy. 

Using blmer to create model?

## Codebook
### Outcome Variables (1-100)

GRID_GTS1_1_GTS1 : Feel at least somewhat worried about climate change

GRID_GTS1_2_GTS1 : Support of Carbon Tax

GRID_GTS1_3_GTS1 : Support of Renewable Energy

GRID_GTS1_4_GTS1 : Support for 100% sourcing electricity from renewable sources

GRID_GTS1_5_GTS1 : Support for New Green Deal

GRID_GTS2_1_GTS2 : Feel at least somewhat worried about climate change (alternate wording)

GRID_GTS2_2_GTS2 : Support of Carbon Tax (alternative wording)

GRID_GTS2_3_GTS2 : Support of Renewable Energy (alternative wording)

GRID_GTS2_4_GTS2 : Support for 100% sourcing electricity from renewable sources (alternative wording)

GRID_GTS2_5_GTS2 : Support for New Green Deal (alternative wording)

### Predictor Variables (subject to change)

USHHI3 : Household Income

USEDU3 : Highest level of education completed

USMAR2 : Marital Status



```{r}
d <- read_csv("full_dataset.csv")


```











